---
title: "Which transmission better for MPG"
author: 
date: 
output: html_document
---
  ### Synopsis
  In this document, we will answer 2 questions below by taking mtcars dataset and using regression models and exploratory data analyses: 
  
  1. Is an automatic or manual transmission better for MPG
2. Quantify the MPG difference between automatic and manual transmissions

### Data Processing

In mtcars dataset, the variable am is for the Transmission (0 = automatic, 1 = manual). 
```{r}
library(datasets);library(car);data(mtcars)
str(mtcars)
```
From plots about mpg vs.am shown in appendix fig-1, we can see that the mean of mpg(miles per gallon) for manual transmission is bigger than mean of automatic, but when considering other variables, the difference is not constant.

```{r}
mtcars$am<-factor(mtcars$am)
fit1<-lm(mpg ~ ., data=mtcars)
summary(fit1)
```
Some p-value are bigger than 0.05, delete the variables whose p-value is biggest. Repeat the step until the all p-values are smaller than 0.05
```{r}
# Adjusted R-squared:  0.8066, delete cyl whose p-value is biggest
fit2<-lm(mpg ~ am + disp + hp + drat + wt + qsec + vs + gear + carb, data=mtcars)
summary(fit2)

# Adjusted R-squared:  0.8153, delete vs whose p-value is biggest
fit3<-lm(mpg ~ am + disp + hp + drat + wt + qsec + gear + carb, data=mtcars)
summary(fit3)

# Adjusted R-squared:  0.823, delete carb whose p-value is biggest
fit4<-lm(mpg ~ am + disp + hp + drat + wt + qsec + gear, data=mtcars)
summary(fit4)

# Adjusted R-squared:  0.8296 , delete gear whose p-value is biggest
fit5<-lm(mpg ~ am + disp + hp + drat + wt + qsec, data=mtcars)
summary(fit5)

# Adjusted R-squared:  0.8347,  delete drat whose p-value is biggest
fit5<-lm(mpg ~ am + disp + hp + wt + qsec, data=mtcars)
summary(fit5)

# Adjusted R-squared:  0.8375, delete disp whose p-value is biggest
fit6<-lm(mpg ~ am + hp + wt + qsec, data=mtcars)
summary(fit6)

# Adjusted R-squared:  0.8368, delete hp whose p-value is biggest
fit7<-lm(mpg ~ am + wt + qsec, data=mtcars)
summary(fit7)
```

all Pr(>F) for model fit7 are less than 0.05 and so model fit7 is selected.  
```{r}
fitn<-lm(mpg~am + wt + qsec, data=mtcars)
```

As the residuals plots about fitn in appendix fig-2, there is any of the patterned appearance. The residuals were independently and (almost) identically distributed with zero mean, and were uncorrelated with the fit and normality.

```{r}
summary(fitn)$coefficients
```

* For am is a two level factor, am==0 is reference, and the Intercept Estimatesv is the mean of the referenct group(am==0) and the other Estimates are the distances of the other groups' means from the reference mean. in average, mpg increase `r round(summary(fitn)$coefficients[2],4)`(E(am==1)) , when changing am from 0 to 1 and holding all the other variables. 
* At the same time, p-values are for tests of whether the groups are different than zero. i.e. H0: beta==0 vs. Ha: beta<>0. In Coefficients table, group(am==1): Pr(>|t|) =`r round(summary(fitn)$coefficients[2,4],4)`<0.05, so we reject H0. i.e. the mpg may be <> 0 when changing am from 0 to 1 and holding other variables.

Considering multivariable with interaction, change am from 0 to 1 while holding other variables constant:

**E[mpgi|am=1,wt=x2,qsec=x3]-[mpgi|am=0,wt=x2,qsec=x3...]=beta_am+beta_wt*x2+beta_qsec*x3**

thus the expected change in per unit change in holding all else constant is not constant. 
Consider t test: **H0:bata_am=`r summary(fitn)$coefficients[2]` vs. Ha:btat_am<>`r summary(fitn)$coefficients[2]`**, calculate the T-confidence interval:
```{r}
m_am1<-summary(fitn)$coefficients[2]
SE<-summary(fitn)$coefficients[2,2]
t<-summary(fitn)$coefficients[2,3]
m_am1+c(-1,1)*SE*t
```
The lower endpoint of T-confidence interval >= 0, so when am change from 0 to 1 and all other variables is held, mpg increase in the interval [`r round(m_am1+c(-1,1)*SE*t,3) `]

### Conclusion
1. Manual transmission is better for MPG than an automatic.
2. The MPG difference between automatic and manual transmissions is not a constant,  which is in [`r round(m_am1+c(-1,1)*SE*t, 3)`] and relies other variables 

### Appendix 
#### Fig-1: mpg ~ am with  cyl as condition
```{r,fig.height=4,fig.width=5}
boxplot(mpg~am, data=mtcars,boxwex = 0.3,main="mpg vs. am")
```
```{r}
coplot(mpg ~ as.factor(am) | as.factor(cyl), data = mtcars, panel = panel.smooth,row=1)
```

#### Fig-2: Residuals plot for fitn1
```{r,fig.height=4,fig.width=5}
par(mfrow = c(1, 1))
rng<-round(c(-1,1)*max(abs(range(resid(fitn)))), 0)
plot(predict(fitn), resid(fitn), main="Resid vs. fit", ylim=rng)
abline(h=0, col="red", lty = 3)
abline(h=4, col="red", lty = 3)
abline(h=-4, col="red", lty = 3)
```
```{r,fig.height=8,fig.width=10}
par(mfrow = c(2, 2)); plot(fitn)
```
