---
title: "Which transmission better for MPG"
author: 
date: 
output: html_document
---
### Synopsis
In this document, we will answer 2 questions below by taking mtcars dataset and using regression models and exploratory data analyses: 

1. Is an automatic or manual transmission better for MPG
2. Quantify the MPG difference between automatic and manual transmissions

### Data Processing

In mtcars dataset, the variable am is for the Transmission (0 = automatic, 1 = manual). 
```{r}
library(datasets);library(car);data(mtcars)
str(mtcars)
```
From plots about mpg vs.am shown in appendix fig-1, we can see that the mean of mpg(miles per gallon) for manual transmission is bigger than mean of automatic, but when considering other variables, the difference is not constant.

Calculate the cor() between mpg and am with other variables. 
```{r}
cor(mtcars$mpg, mtcars[c(2:11)])
cor(as.numeric(mtcars$am), mtcars[c(2:8,10:11)])
```
The cor() result shows that mpg is correlative with all other variables, and am is too. So the am interact with other variables. And they are considered to be added into model together.
```{r}
mtcars$am<-factor(mtcars$am)
fit1<-lm(mpg ~ am, data=mtcars)
fitn1<-lm(mpg ~ ., data=mtcars)
anova(fit1,fitn1)
```
Pr(>F) for model 2 is less than 0.05, and thus it fails to reject F-test. So the variables should be added into model and model fitn is selected.  

As the residuals plots about fitn in appendix fig-2, there is any of the patterned appearance. The residuals were independently and (almost) identically distributed with zero mean, and were uncorrelated with the fit.
But it seems there are outliers. Omit the oulier and redraw the plots shown in appendix fig-3 and item seems better a litter.
```{r}
op<-which(row.names(mtcars)=="Ford Pantera L")
fitn2<-lm(mpg ~ ., data=mtcars[-op,])
```
select fitn2 as model and check it's coefficient:
```{r}
fitn<-fitn2
summary(fitn)$coefficients
```

* For am is a two level factor, am==0 is reference, and the Intercept Estimatesv is the mean of the referenct group(am==0) and the other Estimates are the distances of the other groups' means from the reference mean. in average, mpg increase `r round(summary(fitn)$coefficients[9],4)`(E(am==1)) , when changing am from 0 to 1 and holding all the other variables. 
* At the same time, p-values are for tests of whether the groups are different than zero. i.e. H0: beta==0 vs. Ha: beta<>0. In Coefficients table, group(am==1): Pr(>|t|) =`r round(summary(fitn)$coefficients[9,4],4)`> 0.05, so we fail to reject H0. i.e. the mpg may be 0 when changing am from 0 to 1 and holding other variables.

Considering multivariable with interaction, change am from 0 to 1 while holding other variables constant:

   **E[mpgi|am=1,cyl=x2,disp=x3,...]−E[mpgi|am=0,cyl=x2,disp=x3...]=βam+βcyl * x2+βdisp * x3+...**
   
thus the expected change in per unit change in holding all else constant is not constant. 
Consider t test: **H0:βam=`r summary(fitn)$coefficients[9]` vs. Ha:βam<>`r summary(fitn)$coefficients[9]`**, calculate the T-confidence interval:
```{r}
m_am1<-summary(fitn)$coefficients[9]
SE<-summary(fitn)$coefficients[9,2]
t<-summary(fitn)$coefficients[9,3]
m_am1+c(-1,1)*SE*t
```
The lower endpoint of T-confidence interval >= 0, so when am change from 0 to 1 and all other variables is held, mpg increase in the interval [`r round(m_am1+c(-1,1)*SE*t,3) `]

### Conclusion
1. Manual transmission is better for MPG than an automatic.
2. The MPG difference between automatic and manual transmissions is not a constant,  which is in [`r round(m_am1+c(-1,1)*SE*t, 3)`] and relies other variables 

### Appendix 
#### Fig-1: mpg ~ am with  cyl as condition
```{r,fig.height=4,fig.width=5}
boxplot(mpg~am, data=mtcars,boxwex = 0.3,main="mpg vs. am")
```
```{r}
coplot(mpg ~ as.factor(am) | as.factor(cyl), data = mtcars, panel = panel.smooth,row=1)
````

#### Fig-2: Residuals plot for fitn1
```{r,fig.height=4,fig.width=5}
 par(mfrow = c(1, 1))
 rng<-round(c(-1,1)*max(abs(range(resid(fitn1)))), 0)
 plot(predict(fitn1), resid(fitn1), main="Resid vs. fit", ylim=rng)
 abline(h=0, col="red", lty = 3)
 abline(h=4, col="red", lty = 3)
 abline(h=-4, col="red", lty = 3)
```
```{r,fig.height=8,fig.width=10}
 par(mfrow = c(2, 2)); plot(fitn1)
```

#### Fig-3: Residuals plot for fitn2 after omitting
```{r,fig.height=4,fig.width=5}
 par(mfrow = c(1, 1))
 rng<-round(c(-1,1)*max(abs(range(resid(fitn2)))), 0)
 plot(predict(fitn2), resid(fitn2), main="Resid vs. fit", ylim=rng)
 abline(h=0, col="red", lty = 3)
 abline(h=4, col="red", lty = 3)
 abline(h=-4, col="red", lty = 3)
```
```{r,fig.height=8,fig.width=10}
 par(mfrow = c(2, 2)); plot(fitn2)
```